Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-28 paper/high-assurance-contracts.tex
checkpointing before tackling Section 7

\subsubsection*{Imperative code contracts}
Stream-based evaluation allows values arbitrarily `deep in the past'
to be accessed by use of $\konst{pre}(-)$. However, the programming
languages targeted by SPLAT do not directly support streams.  We are
developing a translation to eliminate such deep temporal accesses by
the introduction of intermediate variables. The essential property the
translation needs to satisfy is a syntactic restriction ---call
it \emph{Pascalish}---that supports easy translation to standard
imperative languages.  The guiding intuition is that a Pascalish
{\Eqs} accesses at most one step in the past; also, accesses to a
past value of a variable are only possible until the variable is
updated. In informal terms, this means that ${\Eqs}\cdot G$ in a code
contract has the following properties:
\begin{itemize}
\item Followed-by only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
\item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
\item $\konst{pre}(v)$ is allowed only before the defining equation $v = e$ (and may occur in $e$).
\end{itemize}
For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be
defined, albeit in a non-Pascalish manner, by
{\small
\begin{lstlisting}[style=agree]
  Fib = 1 -> pre(1 -> Fib + pre Fib)
\end{lstlisting}
}
\noindent or by (again non-Pascalish)
{\small
\begin{lstlisting}[style=agree]
  N = 0 -> 1 + pre(N)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib))
\end{lstlisting}
}
\noindent or by (Pascalish):
{\small
\begin{lstlisting}[style=agree]
  N = 0 -> 1 + pre(N)
  F2 = 42 -> pre(F1)
  F1 = 42 -> pre(Fib)
  Fib = if N <= 1 then 1 else F1 + F2
\end{lstlisting}
}
(Note that the initial values of \verb+F1+ and \verb+F2+ are never
accessed, so 42 could be replaced by any number in their defining
equations.) A Pascalish ${\Eqs} \cdot G$ is straightforward to
implement in conventional programming languages, as we will see in
Section \ref{sec:synthesis}.

@@

As previously discussed, transformations carried out in the BriefCASE
interface have created a cyber-hardened system, protecting against
malicious AI behavior, by inserting a high-assurance filter and
monitor between the AI and the WM (see \figref{fig:hardened}). The
filter protects the WM from malformed data and the monitor protects
the WM from (possibly malicious) spontaneous or delayed responses.
These components were added, one at a time, by
\begin{enumerate}
  \item selecting the connection in the model where the
    component is to appear,
  \item choosing the appropriate transformation, and
  \item specifying the relevant filtering or monitoring \emph{policy} (behavior).
\end{enumerate}

@@

\subsubsection{Monitor specification}
BriefCASE generates everything but the
\konst{eq}-statements in \linesref{line:mon-assume}{line:mon-policy} of
\figref{fig:monitor}.
The
\texttt{alert} variable defined on \lineref{line:mon-alert} depends
not just on the policy but also on configuration details supplied when
the component is specified.  At configuration
time, \texttt{is\_latched} has been set to \konst{true}, thus making
the alert persistent, meaning that once the alert is raised, it is
always raised; otherwise, it is the complement of the policy value in
the current step.

@@

As shown in \figref{fig:hardened-certificate}, the cyber-hardened
system passes AGREE verification.  That means that the high-assurance
components protect the WM from the untrusted AI generating malformed
data, spontaneous responses, or delayed responses.  Synthesis must
generate precise implementations of the high-assurance component
specifications for the AGREE verification result to apply to the
deployed system.  Precise in this context means that the
implementations match the input-to-output relations defined in the
specifications.

>>>>

\subsubsection*{Imperative code contracts.}
Stream-based evaluation allows values arbitrarily `deep in the past'
to be accessed by use of $\konst{pre}(-)$. However, the programming
languages targeted by SPLAT do not directly support streams.  We are
developing a translation to eliminate such deep temporal accesses by
the introduction of intermediate variables. The essential property the
translation needs to satisfy is a syntactic restriction ---call
it \emph{Pascalish}---that supports easy translation to standard
imperative languages.  The guiding intuition is that a Pascalish
{\Eqs} accesses at most one step in the past; also, accesses to a
past value of a variable are only possible until the variable is
updated. In informal terms, this means that ${\Eqs}\cdot G$ in a code
contract has the following properties:
\begin{itemize}
\item Followed-by only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
\item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
\item $\konst{pre}(v)$ is allowed only before the defining equation $v = e$ (and may occur in $e$).
\end{itemize}
For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be
defined, albeit in a non-Pascalish manner, by
{\small
\begin{lstlisting}[style=agree]
  Fib = 1 -> pre(1 -> Fib + pre Fib)
\end{lstlisting}
}
\noindent or by (again non-Pascalish)
{\small
\begin{lstlisting}[style=agree]
  N = 0 -> 1 + pre(N)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib))
\end{lstlisting}
}
\noindent or by (Pascalish):
{\small
\begin{lstlisting}[style=agree]
  N = 0 -> 1 + pre(N)
  F2 = 42 -> pre(F1)
  F1 = 42 -> pre(Fib)
  Fib = if N <= 1 then 1 else F1 + F2
\end{lstlisting}
}
(Note that the initial values of \verb+F1+ and \verb+F2+ are never
accessed, so 42 could be replaced by any number in their defining
equations.) A Pascalish ${\Eqs} \cdot G$ is straightforward to
implement in conventional programming languages, as we will see in
Section \ref{sec:synthesis}.

@@

As previously discussed, transformations carried out in the BriefCASE
interface have created a cyber-hardened system, protecting against
malicious AI behavior by the insertion of a high-assurance filter and
monitor between the AI and the WM (see \figref{fig:hardened}). The
filter protects the WM from malformed data and the monitor protects
the WM from (possibly malicious) spontaneous or delayed responses.
These components were added, one at a time, by
\begin{enumerate}
  \item selecting the connection in the model where the
    component is to appear,
  \item choosing the appropriate transformation, and
  \item specifying the relevant filtering or monitoring \emph{policy} (behavior).
\end{enumerate}

@@

\subsubsection{Monitor specification}
BriefCASE generates everything but the
\konst{eq}-statements in \linesref{line:mon-assume}{line:mon-policy} of
\figref{fig:monitor}.

The \texttt{alert} variable defined on \lineref{line:mon-alert}
(distinct from the \texttt{Alert} port) depends not just on the policy
but also on configuration details supplied when the component is
specified.  At configuration time, \texttt{is\_latched} has been set
to \konst{true}, thus making the alert persistent, meaning that once
the alert is raised, it is always raised; otherwise, it is the
complement of the policy value in the current step.

@@

As shown in \figref{fig:hardened-certificate}, the cyber-hardened
system passes AGREE verification.  That means that the high-assurance
components protect the WM from the untrusted AI generating malformed
data, spontaneous responses, or delayed responses.  Synthesis must
generate faithful implementations of the high-assurance component
specifications for the AGREE verification result to apply to the
deployed system.

%%   Precise in this context means that the
%% implementations match the input-to-output relations defined in the
%% specifications.
